Introducing statistics with \\texttt{intRo}

### 2014-12-05 student-paper-competition-2014/paper.Rmd
remove some comments

---
title: "Introducing statistics with \\texttt{intRo}"
author: "Eric Hare & Andee Kaplan"
date: "December 14, 2014"
output:
  pdf_document:
    includes:
      in_header: preamble.tex
    number_sections: yes
bibliography: refs.bib
---

\begin{abstract}
\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}. \texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}. Within this paper we introduce the application and explore the design decisions underlying \texttt{intRo} as well as highlight some challenges and advantages of reactive programming.
\end{abstract}

\newpage

@@

The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure~\ref{fig:app_creation_modules}. The key driver to populating \texttt{server.R} and \texttt{ui.R} is the file \texttt{modules.txt}, which defines the placement of each module. The interface is then created with the following statement.

@@

In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability. At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{Shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability. For a tangible example, consider the graphical summaries module. We only allow variables of a type consistent with the selected plot to be displayed. This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application. On the flip side of this, \texttt{Shiny} allows much higher flexibility. For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}. \ak{[Maybe add a figure here?, including example code for dynamically loading the ui]} However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability. One current example the slider element. This element allows for fixed width steps from its minimum to its maximum. The JavaScript library being utilized in \texttt{Shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript. This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.

>>>>

---
title: "Introducing statistics with \\texttt{intRo}"
author: "Eric Hare & Andee Kaplan"
date: "December 14, 2014"
output:
  pdf_document:
    includes:
      in_header: preamble.tex
    number_sections: yes
bibliography: refs.bib
---
\thispagestyle{empty}
\begin{abstract}
\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}. \texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}. Within this paper we introduce the application and explore the design decisions underlying \texttt{intRo} as well as highlight some challenges and advantages of reactive programming.
\end{abstract}

\newpage
\setcounter{page}{1}

@@

The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}. The key driver to populating \texttt{server.R} and \texttt{ui.R} is the file \texttt{modules.txt}, which defines the placement of each module. The interface is then created with the following statement.

@@

In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability. At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{Shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability. For a tangible example, consider the graphical summaries module. We only allow variables of a type consistent with the selected plot to be displayed. This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application. On the flip side of this, \texttt{Shiny} allows much higher flexibility. For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}, as detailed in section \ref{sec:extensible}. However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability. One current example the slider element. This element allows for fixed width steps from its minimum to its maximum. The JavaScript library being utilized in \texttt{Shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript. This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.
